---
title: 'Code source: connecting SQL to R'
author: "Firza Riany"
date: "16/01/2021"
output: html_document
---

## Establish a connection

```{r message=FALSE, warning=FALSE}
library(DBI)
library(dbplyr)
library(dplyr)
library(RPostgreSQL)

rental = DBI::dbConnect(odbc::odbc(),
                        driver = "PostgreSQL ANSI(x64)",
                        database = "dvdrental_2",
                        port = 5433,
                        host = "localhost",
                        UID = rstudioapi::askForPassword("Database user"),
                        PWD = rstudioapi::askForPassword("Database password"))
# username = postgres
# password = R4hasiabanget
```

```{r connex 2}
mammals <- DBI::dbConnect(RSQLite::SQLite(), "data_raw/portal_mammals.sqlite")

rental_2 = DBI::dbConnect(RPostgreSQL::PostgreSQL(), "dvdrental")
```

## Simple queries

To see the tables that the database contains

```{r check tables}
src_dbi(rental)
```

You can also check the tables from the connection tab in the upper right box of your RStudio. It basically serves a similar function as when you open the schema tab in SQL.

## Navigating through the database using SQL queries in R

If you want to navigate through the database using SQL queries, you can use function tbl() followed by the SQL queries. This function is merely asking R to send queries to SQL, so that SQL can do the relevant work with the database.

```{r select everything}
# Select everything from the film table

tbl(rental, sql("SELECT * FROM film"))
```

```{r select count}
# How many rating groups are listed in the film table

tbl(rental, sql("SELECT COUNT(DISTINCT(rating)) FROM film"))

```

```{r select distinct}
# Listing all the rating groups present in the film table

tbl(rental, sql("SELECT DISTINCT(rating) FROM film"))
```

```{r select where}
# Selecting the film titles where rating groups are PG and PG-13

tbl(rental, sql("SELECT title, length, replacement_cost, rental_rate, rating
                FROM film
                WHERE rating = 'PG-13' OR rating = 'PG'
                ORDER BY rental_rate ASC"))
```

Besides using the tbl() function to navigate a database, you can also use dplyr verbs. Here are the alternative syntax in dplyr for each of the SQL queries in the above.

The same operation can be done using dplyr syntax. For example, to select several variables (or column) from the table \*"surveys"\*, I can first get the table using tbl() function then I select the variables by passing on the names of the variables in the arguments of select().

```{r dplyr select}
# Selecting everything from the film table

film_table = tbl(rental, "film")

film_table %>%
  select(film_id, title, rental_rate, replacement_cost, rating)
```

```{r dplyr select count}
# How many rating groups are listed in the film table

film_table %>% 
  distinct(rating) %>%
  count()
```

```{r dplyr select distinct}
# Listing all the rating groups present in the film table

film_table %>%
  distinct(rating)
```

```{r select filter dplyr}
subset_film = 
  film_table %>%
  filter(rating == "PG-13" | rating == "PG") %>%
  select(title, length, replacement_cost, rental_rate, rating) %>%
  arrange(rental_rate) %>%
  collect()

dim(subset_film)

class(subset_film)
```

```{r sans collect}
subset_film_0 = 
  film_table %>%
  filter(rating == "PG-13" | rating == "PG") %>%
  select(title, length, replacement_cost, rental_rate, rating) %>%
  arrange(rental_rate)

dim(subset_film_0)

class(subset_film_0)
```

```{r question_1}
# This is to find the most rented films
dbGetQuery(rental,
           "SELECT film.title, COUNT(inventory.film_id) AS counter
           FROM inventory
           INNER JOIN rental
            ON inventory.inventory_id = rental.inventory_id
           INNER JOIN film
            ON inventory.film_id= film.film_id
           GROUP BY film.film_id
           ORDER BY counter DESC
           LIMIT 5")
```

To find the most rented films

I need to get the title of the films that have been rented the most

To do that, I need to check how many rentals are made for each film. This information can be found in the *rental* table. Naturally, we can just join *rental* table and *film* table by the same identifier. However, we can't find the same identifier between *rental* and *film* table.

```{r film table}
film_db = tbl(rental, "film")  # getting the table

head(film_db, n = 5)
```

```{r rental table}
rental_db = tbl(rental, "rental")

head(rental_db, n = 5)
```

So, there are no similar identifier between *rental* and *film* table. However, there is inventory_id in *rental* table which can also be found in *inventory* table. When we check *inventory* table, turned out there is film_id, the identifier to find film title in the *film* table.

```{r inventory table}
inventory_db = tbl(rental, "inventory")

head(inventory_db, n = 5)
```

In a way, *inventory* table can be the bridge between *rental* and *film* tables to find the total rentals for each film title.

What we need to do now is we can inner join *rental* and *inventory* tables by inventory_id

```{r join rental x invent}
join_tbl = inner_join(rental_db, inventory_db, by = "inventory_id")

join_tbl
```

Then to get the total rentals of film, we count the total rentals by film_id. In a way, we aggregate the total rentals for each film_id. In SQL, we use COUNT(film_id) ... GROUP BY(film_id).

So, after joining *rental* and *inventory* tables (stored in join_tbl), I group the film_id, then count how many film_id appears in the join_tbl. (stored in **count** variable) To get the film title, I join the results with film_db. To find the most rented films, I order the results based on **count** variable in an descending order. For the final touch, I select only variable **title, count, replacement_cost, and rating** from film_db. Then I will have to store this in my R's environment by passing collect() function at the end of the syntax.

```{r most_rented}
most_rented = 
  join_tbl %>% 
  group_by(film_id) %>%
  summarise(count = n()) %>%
  inner_join(film_db, by = "film_id") %>%
  arrange(desc(count)) %>%
  select(title, count, replacement_cost, rating) %>%
  collect()

most_rented
```

To get the top five, I will use the function head() with argument n = 5.

```{r top_five}
top_five = head(most_rented, n = 5)

top_five
```

In SQL, you should pass the following code to get the same results

```{r SQL top_five}
tbl(rental, sql("SELECT title, COUNT(inventory.film_id) AS counter
           FROM inventory
           INNER JOIN rental
            ON inventory.inventory_id = rental.inventory_id
           INNER JOIN film
            ON inventory.film_id= film.film_id
           GROUP BY film.film_id
           ORDER BY counter DESC
           LIMIT 5"))
```

The pipeline of the syntax is basically the same. So I create nested inner join. Firstly between *rental* and *inventory* tables on inventory_id, which is the identifier presents in both tables. And secondly between the results of the previous join with *film* table on film_id which is the identifier presents in *film* and *inventory* table.

And I count the multiple occurrence of film_id grouped by film_id and store the result in **count** variable. Then I order the result in a descending order and return only the top five rows.
